Case study validations of automatic bullet matching\tnoteref{t1}\tnoteref{t2}

### 2019-11-10 index.Rnw
Clean up figure 2

<<echo=FALSE, results = 'hide', warning = FALSE, message = FALSE>>=
features <- read.csv("data/fsi-features.csv.gz")
features <- features %>% mutate(
  bullet1 = gsub("ullet ", "", bullet1),
  bullet2 = gsub("ullet ", "", bullet2)
)
b12 <- features %>% filter(
  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
)
p1 <- b12 %>%
  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
  geom_tile() + ggtitle("Same barrel") +
  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
  theme_bw() +
  xlab("Bullet 2") + ylab("Bullet 1") +
  theme(legend.position = "bottom") + coord_equal()

bab <- features %>% filter(
  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
)
p2 <- bab %>%
  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
  geom_tile() + ggtitle("Different barrel") +
  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
  theme_bw() +
  xlab("Bullet 2") + ylab("Bullet 1") +
  theme(legend.position = "bottom") + coord_equal()
@

@@

<<b2b, echo = FALSE, fig.width = 5.4, fig.height = 3, out.width = ".6\\textwidth", fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons. On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels).">>=
grid.arrange(p1, p2, ncol = 2, padding = 2)
@

When imaging bullets using a confocal light microscope, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence (Other imaging systems, such as BulletTrax 3D\footnote{https://www.ultra-forensictechnology.com/en/our-products/ballistic-identification/bullettrax/} or BalScan\footnote{https://www.forensic.cz/en/products/balscan}, may capture lower-resolution scans of an entire bullet at one time). The order in which scans are acquired is kept as meta-information. Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel. For all of the bullets considered here the number of lands $p$ is 6. A match between  a pair of lands from two bullets therefore results in an expected additional $p-1$ matches between pairs of lands. These lands are also expected to be in a sequence, i.e. if there is a match between lands $\ell_i$ on bullet 1 and $\ell_j$ on bullet 2, we also expect lands $\ell_{i\oplus s}$ and $\ell_{j\oplus s}$ to match for all integers $s$, where $\oplus$ is defined as $a \oplus b \equiv \left((a + b - 1)\mod p\right) + 1$. This relationship gives rise to the sequence average maximum (SAM) to quantify a bullet-to-bullet match.

>>>>

<<echo=FALSE, results = 'hide', warning = FALSE, message = FALSE>>=
# features <- read.csv("data/fsi-features.csv.gz")
# features <- features %>% mutate(
#   bullet1 = gsub("ullet ", "", bullet1),
#   bullet2 = gsub("ullet ", "", bullet2)
# )
# b12 <- features %>% filter(
#   group1 == "G1", barrel1 == "KA", bullet1 == "B1",
#   group2 == "G1", barrel2 == "KA", bullet2 == "B2"
# )
# 
# p1 <- b12 %>%
#   ggplot(aes(x = land1, y = land2, fill = rfscore)) +
#   geom_tile() + ggtitle("Same barrel") +
#   scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
#   theme_bw() +
#   xlab("Bullet 2") + ylab("Bullet 1") +
#   theme(legend.position = "bottom") + coord_equal()
# 
# bab <- features %>% filter(
#   group1 == "G1", barrel1 == "KA", bullet1 == "B1",
#   group2 == "G1", barrel2 == "KB", bullet2 == "B1"
# )
# p2 <- bab %>%
#   ggplot(aes(x = land1, y = land2, fill = rfscore)) +
#   geom_tile() + ggtitle("Different barrel") +
#   scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
#   theme_bw() +
#   xlab("Bullet 2") + ylab("Bullet 1") +
#   theme(legend.position = "bottom") + coord_equal()

l2lscores <- bind_rows(
  mutate(b12, type = "Same barrel"),
  mutate(bab, type = "Different barrel")
) %>% 
  mutate(type = factor(type, levels = c("Same barrel", "Different barrel"), ordered = T))

@@

<<b2b, echo = FALSE, fig.width = 5.4, fig.height = 3, out.width = ".6\\textwidth", fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons. On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels).">>=
# grid.arrange(p1, p2, ncol = 2, padding = 2)
ggplot(data = l2lscores, aes(x = land1, y = land2, fill = rfscore)) +
  geom_tile() + 
  scale_fill_gradient2("Random Forest (RF) score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
  theme_bw() +
  xlab("Bullet 2") + ylab("Bullet 1") +
  theme(legend.position = "bottom") + coord_equal() + facet_wrap(~type)
@

While some imaging systems, such as BulletTrax 3D\footnote{https://www.ultra-forensictechnology.com/en/our-products/ballistic-identification/bullettrax/} or BalScan\footnote{https://www.forensic.cz/en/products/balscan}, may capture lower-resolution scans of an entire bullet at one time, confocal light microscopes do not have 360$^\circ$ capability. Thus, when imaging bullets using a confocal light microscope, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence. The order in which scans are acquired is kept as meta-information.

Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel. For all of the bullets considered here the number of lands $p$ is 6. A match between  a pair of lands from two bullets therefore results in an expected additional $p-1$ matches between pairs of lands. These lands are also expected to be in a sequence, i.e. if there is a match between lands $\ell_i$ on bullet 1 and $\ell_j$ on bullet 2, we also expect lands $\ell_{i\oplus s}$ and $\ell_{j\oplus s}$ to match for all integers $s$, where $\oplus$ is defined as $a \oplus b \equiv \left((a + b - 1)\mod p\right) + 1$. This relationship gives rise to the sequence average maximum (SAM) to quantify a bullet-to-bullet match.
